\pkg{depmixS4} : An \proglang{R}-package for hidden Markov models

### 2010-03-03 papers/jss/dpx4Rev.Rnw
Jss paper updates, removed .bib files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@374 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\paragraph{Starting values} Note also that start values for the
transition parameters are provided in this call using the
\code{trstart} argument.  At this time, the package does not provide
automatic starting values.  Starting values for parameters can be
provided using three arguments: \code{instart} for the parameters
relating to the prior probabilities, \coce{trstart}, relating the
transition models, and \code{respstart} for the parameters of the
response models.  Note that the starting values for the initial and
transition models as well as of the multinomial logit response models
are interpreted as {\em probabilities}, and internally converted to
multinomial logit parameters.

@@

\subsection{Covariates on transition parameters\label{sec:trans}}

@@

\subsection{Adding covariates on the prior probabilities\label{sec:prior}}

@@

The balance scale taks is a famous task for testing cognitive
strategies developed by Piaget \citep{Siegler1981}.
Figure~\ref{fig:balance} provides an example of a balance scale item.
Participants' task is to say to which side the balance will tip when
released, or alternatively, whether it will stay in balance.  The item
shown in Figure~\ref{fig:balance} is a so-callled distance item: the
number of weights placed on each side is equal, and only the distance
of the weights to the fulcrum differs between each side.

\begin{figure}[htbp]
\begin{center}
	\includegraphic{graphs:baldist.pdf}
	\caption{Balance scale item; this is a so-called distance item.}
	\label{fig:balance}  
\end{center}
\end{figure}

Children in the lower grades of primary school are known to ignore the
distance dimension, and base their answer only on the number of
weights on each side.  Hence, they would typically provide the wrong
answer to these distance items.  Slightly older children do take
distance into account when responding to balance scale items, but they
only do so when the number of weights is equal on each side.  These
two strategies that children employ are known a Rule I and Rule II.
Other strategies can be teased apart by administering different items.
The \code{balance} data set that we analyse here consists of 4
distance items on a balance scale task administered to 779
participants ranging from 5 to 19 years of age.  The full set of items
consisted of 25 items; other items in the test are used to detect
other strategies that children (and young adults) employ in solving
balance scale items \citep[see][for details]{Jansen2002}.

Similarly to the transition matrix, covariates on the prior
probabilities of the latent states (or classes in this case), are
defined by using a one-sided formula:
<<>>=
data(balance)
balance$age <- balance$age-5
set.seed(1)
mod <- mix(list(d1~1,d2~1,d3~1,d4~1), data=balance, nstates=2,
    family=list(multinomial(), multinomial(), multinomial(),
    multinomial()), respstart=c(rep(c(0.9,0.1),4),rep(c(0.1,0.9),4)), 
    prior=~age, initdata=balance)
fm <- fit(mod)
@

@@

@@

%
% add more explanation about the parameters, possibly add a plot of 
% the class membership as function of age
%

Hence at young ages, children have a high probability of incorrect 
responding in class~1, whereas the prior probability for class~2 
increases with age.

@@

To see which parameters are fixed (by default only baseline parameters
in the multinomial logistic models for the transition models and the
initial state probabilities model):
<<>>=
setpars(mod, getpars(mod,which="fixed"))

@
When fitting constraints it is useful to have good starting values 
for the parameters and hence we first fit the following model without
constraints:
<<>>=
trst <- c(0.9,0.1,0,0,0.1,0.9,0,0)
mod <- depmix(list(rt~1,corr~1),data=speed,transition=~Pacc,
	  nstates=2,family=list(gaussian(),multinomial()),
	  trstart=trst,inst=c(.999,0.001))
fm1 <- fit(mod)
@

After this, we use the fitted values from this model to constrain the
regression coefficients on the transition matrix (parameters numbers~6
and~10):
<<>>=
# start with fixed and free parameters
conpat <- c(0,1,rep(c(0,1),4),1,1,0,1,1,1,0,1)
# constrain the beta's on the transition parameters to be equal
conpat[6] <- conpat[10] <- 2
fm2 <- fit(fm1,equal=conpat)
@

>>>>

\paragraph{Starting values} Note also that start values for the
transition parameters are provided in this call using the
\code{trstart} argument.  At this time, the package does not provide
automatic starting values.  Starting values for parameters can be
provided using three arguments: \code{instart} for the parameters
relating to the prior probabilities, \code{trstart}, relating the
transition models, and \code{respstart} for the parameters of the
response models.  Note that the starting values for the initial and
transition models as well as of the multinomial logit response models
are interpreted as {\em probabilities}, and internally converted to
multinomial logit parameters.

@@

\subsection{Covariates on transition parameters}

@@

\subsection{Adding covariates on the prior probabilities}

@@

The balance scale taks is a famous task for testing cognitive
strategies developed by Jean Piaget \citep[see]{Siegler1981}.
Figure~\ref{fig:balance} provides an example of a balance scale item.
Participants' task is to say to which side the balance will tip when
released, or alternatively, whether it will stay in balance.  The item
shown in Figure~\ref{fig:balance} is a so-callled distance item: the
number of weights placed on each side is equal, and only the distance
of the weights to the fulcrum differs between each side.

\begin{figure}[htbp]
\begin{center}
	\includegraphics{graphs/baldist.pdf}
	\caption{Balance scale item; this is a so-called distance item.}
	\label{fig:balance}  
\end{center}
\end{figure}

Children in the lower grades of primary school are known to ignore the
distance dimension, and base their answer only on the number of
weights on each side.  Hence, they would typically provide the wrong
answer to these distance items.  Slightly older children do take
distance into account when responding to balance scale items, but they
only do so when the number of weights is equal on each side.  These
two strategies that children employ are known as Rule I and Rule II.
Other strategies can be teased apart by administering different items.
The \code{balance} data set that we analyse here consists of 4
distance items on a balance scale task administered to 779
participants ranging from 5 to 19 years of age.  The full set of items
consisted of 25 items; other items in the test are used to detect
other strategies that children and young adults employ in solving
balance scale items \citep[see][for details]{Jansen2002}.

In the following model, age is included as covariate on class
membership to test whether, with age, children apply more complex
rules in solving balance scale items.  Similarly to the transition
matrix, covariates on the prior probabilities of the latent states (or
classes in this case), are defined by using a one-sided formula
\code{prior=~age}: 
<<>>= 
data(balance)
set.seed(1)
mod <- mix(list(d1~1,d2~1,d3~1,d4~1), data=balance, nstates=3,
	family=list(multinomial("identity"), multinomial("identity"),
	multinomial("identity"), multinomial("identity")),
	respstart=runif(24), prior=~age, initdata=balance)
fm <- fit(mod) 
fm
@

@@

As can be seen from the print of the fitted model above, the BIC for 
this model equals 1941.6. The similarly defined 2-class model for 
these data has a BIC of 1969.2, and the 4-class model has BIC equal to
1950.4. Hence, the 3-class seems to be adequate for describing these 
data.

@@

The intercept values of the multinomial logistic parameters provide
the prior probabilities of each class when the covariate has value
zero (note that in this case the value zero does not have much
significance, scaling and/or centering of covariates may be useful in
such cases).  The summary function prints these values.  As can be
seen from those values, at age zero, the prior probability is
overwhelmingly at the second class.  Inspection of the response
parameters reveals that class~2 is associated with incorrect
responding, whereas class~1 is associated with correct responding;
class~3 is an intermediate class with guessing behavior.
Figure~\ref{fig:balprior} depicts the prior class probabilities as
function of age based on the fitted parameters

\begin{figure}[htbp]
\begin{center}
<<fig=TRUE,echo=FALSE,height=5,width=7,eps=FALSE>>=
x <- mlogit(base=1)
coeff <- coefficients(fm@prior@parameters)

pr1 <- function(y) {sapply(y, function(z) {x$linkinv(c(t(coeff)%*%c(1,z)), base=1)[1]})}
pr2 <- function(y) {sapply(y, function(z) {x$linkinv(c(t(coeff)%*%c(1,z)), base=1)[2]})}
pr3 <- function(y) {sapply(y, function(z) {x$linkinv(c(t(coeff)%*%c(1,z)), base=1)[3]})}

plot(pr1,min(balance$age),max(balance$age),lty=1,ylim=c(0,1),
main="Prior probabilities by age, balance scale data", xlab="age", ylab="Pr")
plot(pr2,min(balance$age),max(balance$age),add=T,lty=2)
plot(pr3,min(balance$age),max(balance$age),add=T,lty=3)

legend("right",legend=c("Class 1 (correct)","Class 2 (incorrect)","Class 3 (guess)"),lty=1:3,inset=c(0.1,0))
@
	\caption{Class probabilities as a function of age.}
	\label{fig:balprior}  
\end{center}
\end{figure}

As can be seen from Figure~\ref{fig:balprior}, at younger ages 
children predominantly apply Rule I, the wrong strategy for these 
items. According to the model, approximately 90 \% of children at age
5 apply Rule I. The remaining 10 \% are evenly split among the 2 
other classes. At age 19, almost all participants belong to class~1. 
Interestingly, prior probability of the 'guess' class 2, first 
increases with age, and then decreases again. This suggests that 
children pass through a phase in which they are uncertain or possibly 
switch between applying different strategies.

@@

To see which parameters are fixed (by default only baseline parameters
in the multinomial logistic models for the transition models and the
initial state probabilities model):
<<>>=
setpars(mod, getpars(mod,which="fixed"))
@
When fitting constraints it is useful to have good starting values 
for the parameters and hence we first fit the following model without
constraints:
<<>>=
trst <- c(0.9,0.1,0,0,0.1,0.9,0,0)
mod <- depmix(list(rt~1,corr~1),data=speed,transition=~Pacc,
	  nstates=2,family=list(gaussian(),multinomial()),
	  trstart=trst,inst=c(.999,0.001))
fm1 <- fit(mod)
@

After this, we use the fitted values from this model to constrain the
regression coefficients on the transition matrix (parameters number~6
and~10):
<<>>=
# start with fixed and free parameters
conpat <- c(0,1,rep(c(0,1),4),1,1,0,1,1,1,0,1)
# constrain the beta's on the transition parameters to be equal
conpat[6] <- conpat[10] <- 2
fm2 <- fit(fm1,equal=conpat)
@

### 2010-03-03 papers/jss/dpx4Rev.tex
Jss paper updates, removed .bib files
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/depmix@374 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

\paragraph{Starting values} Note also that start values for the
transition parameters are provided in this call using the
\code{trstart} argument.  At this time, the package does not provide
automatic starting values.  Starting values for parameters can be
provided using three arguments: \code{instart} for the parameters
relating to the prior probabilities, \coce{trstart}, relating the
transition models, and \code{respstart} for the parameters of the
response models.  Note that the starting values for the initial and
transition models as well as of the multinomial logit response models
are interpreted as {\em probabilities}, and internally converted to
multinomial logit parameters.

\paragraph{Fitting the model and printing results} The \code{depmix}
function returns an object of class \code{depmix} which contains the
model specification, and not a fitted model.  Hence, the model needs
to be fitted by calling \code{fit}:
\begin{Schunk}
\begin{Sinput}
> fm <- fit(mod)
\end{Sinput}
\begin{Soutput}
iteration 0 logLik: -305.3302 
iteration 5 logLik: -305.3251 
iteration 10 logLik: -305.3133 
iteration 15 logLik: -305.2824 
iteration 20 logLik: -305.1948 
iteration 25 logLik: -304.8879 
iteration 30 logLik: -302.2484 
iteration 35 logLik: -109.1782 
iteration 40 logLik: -84.3722 
iteration 45 logLik: -84.34183 
iteration 50 logLik: -84.34175 
iteration 50 logLik: -84.34175 
\end{Soutput}
\end{Schunk}

The \code{fit}-function returns an object of class
\code{depmix.fitted} which extends the \code{depmix} class, adding
convergence information (and information about constraints if these
were applied, see below).  The class has \code{print} and
\code{summary} methods to see the results.  The \code{print} method
provides summary information on convergence, the log-likelihood and
the AIC and BIC values:
\begin{Schunk}
\begin{Sinput}
> fm
\end{Sinput}
\begin{Soutput}
Convergence info: Log likelihood converged to within tol. (relative change crit.) 
'log Lik.' -84.34175 (df=7)
AIC:  182.6835 
BIC:  211.275 
\end{Soutput}
\end{Schunk}

@@

The \code{summary} method of \code{fit}ted models provides the
parameter estimates, first for the prior probabilities model, second
for the transition models, and third for the response models.
\begin{Schunk}
\begin{Sinput}
> summary(fm)
\end{Sinput}
\begin{Soutput}
Initial state probabilties model 
Model of type multinomial, formula: ~1
Coefficients: 
     [,1]      [,2]
[1,]    0 -11.25688
Probalities at zero values of the covariates.
0.999987 1.291797e-05

Transition model for state (component) 1 
Model of type multinomial, formula: ~1
Coefficients: 
[1]  0.000000 -2.392455
Probalities at zero values of the covariates.
0.9162501 0.08374986

Transition model for state (component) 2 
Model of type multinomial, formula: ~1
Coefficients: 
[1] 0.000000 2.139255
Probalities at zero values of the covariates.
0.1053396 0.8946604

@@

Response model for response 1 
Model of type gaussian, formula: rt ~ 1
Coefficients: 
[1] 6.385492
sd  0.2439376

@@

Response model for response 1 
Model of type gaussian, formula: rt ~ 1
Coefficients: 
[1] 5.511151
sd  0.1926063 
\end{Soutput}
\end{Schunk}

@@

\begin{figure}[htbp]
\begin{center}
	\includegraphic{graphs:baldist.pdf}
	\caption{Balance scale item; this is a so-called distance item.}
	\label{fig:balance}  
\end{center}
\end{figure}

@@

    1 fx=   0.000000e+00 upsi=  6.7e+00 b2n= -1.0e+00 umi=  0.0e+00 nr   1 si-1
    2 fx=   2.640133e+02 upsi=  0.0e+00 b2n=  2.1e+01 umi=  0.0e+00 nr   1 si-1
    3 fx=   2.636388e+02 upsi=  0.0e+00 b2n=  5.4e+01 umi=  0.0e+00 nr   1 si-1
    4 fx=   2.587645e+02 upsi=  0.0e+00 b2n=  2.7e+02 umi=  0.0e+00 nr   1 si-1
    5 fx=   2.528471e+02 upsi=  0.0e+00 b2n=  1.2e+02 umi=  0.0e+00 nr   1 si-1
    6 fx=   2.523984e+02 upsi=  0.0e+00 b2n=  9.6e+01 umi=  0.0e+00 nr   1 si-1
    7 fx=   2.523907e+02 upsi=  0.0e+00 b2n=  9.5e+01 umi=  0.0e+00 nr   1 si-1
    8 fx=   2.522057e+02 upsi=  0.0e+00 b2n=  9.2e+01 umi=  0.0e+00 nr   1 si-1
    9 fx=   2.515501e+02 upsi=  0.0e+00 b2n=  5.8e+01 umi=  0.0e+00 nr   1 si-1
   10 fx=   2.510856e+02 upsi=  0.0e+00 b2n=  5.2e+01 umi=  0.0e+00 nr   1 si-1
   11 fx=   2.507418e+02 upsi=  0.0e+00 b2n=  7.2e+01 umi=  0.0e+00 nr   1 si-1
   12 fx=   2.503033e+02 upsi=  0.0e+00 b2n=  1.9e+01 umi=  0.0e+00 nr   1 si-1
   13 fx=   2.502725e+02 upsi=  0.0e+00 b2n=  9.8e-01 umi=  0.0e+00 nr   1 si-1
   14 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.5e-01 umi=  0.0e+00 nr   1 si-1
   15 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  3.1e-02 umi=  0.0e+00 nr   1 si-1
   16 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  4.1e-02 umi=  0.0e+00 nr   1 si-1
   17 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  6.3e-02 umi=  0.0e+00 nr   1 si-1
   18 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  2.9e-02 umi=  0.0e+00 nr   1 si-1
   19 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  4.1e-03 umi=  0.0e+00 nr   1 si-1
   20 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.4e-04 umi=  0.0e+00 nr   1 si-1
   21 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  4.3e-03 umi=  0.0e+00 nr   1 si-1
   22 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  9.4e-02 umi=  0.0e+00 nr   1 si-1
   23 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.1e-01 umi=  0.0e+00 nr   1 si-1
   24 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.1e-01 umi=  0.0e+00 nr   1 si-1
   25 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  2.7e-02 umi=  0.0e+00 nr   1 si-1
   26 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  4.6e-02 umi=  0.0e+00 nr   1 si-1
   27 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  2.0e-03 umi=  0.0e+00 nr   1 si-1
   28 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  2.0e-02 umi=  0.0e+00 nr   1 si-1
   29 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  4.9e-02 umi=  0.0e+00 nr   1 si-1
   30 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   31 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   32 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   33 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   34 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   35 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   36 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.1e-01 umi=  0.0e+00 nr   1 si-1
   37 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.1e-01 umi=  0.0e+00 nr   1 si-1
   38 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.5e-02 umi=  0.0e+00 nr   1 si-1

@@

 termination reason:
 KT-conditions satisfied, no further correction computed
 evaluations of f                          149
 evaluations of grad f                      38
 evaluations of constraints                150
 evaluations of grads of constraints         0
 final scaling of objective           2.578752e+00
 norm of grad(f)                      5.469760e-01
 lagrangian violation                 3.850303e-06
 feasibility violation                0.000000e+00
 dual feasibility violation           0.000000e+00
 optimizer runtime sec's              1.093723e+02

@@

The constructor method return an object of class \code{exgaus}, and is
defined as follows:
\begin{Schunk}
\begin{Sinput}
> setGeneric("exgaus", function(y, pstart = NULL, fixed = NULL, 
+     ...) standardGeneric("exgaus"))
\end{Sinput}
\begin{Soutput}
[1] "exgaus"
\end{Soutput}
\begin{Sinput}
> require(gamlss)
\end{Sinput}
\begin{Soutput}
 **********   GAMLSS Version 3.1-0 ********** 
For more on GAMLSS look at http://www.gamlss.com/ 
Type gamlssNews() to see new features/changes/bug fixes.
\end{Soutput}
\begin{Sinput}
> require(gamlss.dist)
> setMethod("exgaus", signature(y = "ANY"), function(y, pstart = NULL, 
+     fixed = NULL, ...) {
+     y <- matrix(y, length(y))
+     x <- matrix(1)
+     parameters <- list()
+     npar <- 3
+     if (is.null(fixed)) 
+         fixed <- as.logical(rep(0, npar))
+     if (!is.null(pstart)) {
+         if (length(pstart) != npar) 
+             stop("length of 'pstart' must be ", npar)
+         parameters$mu <- pstart[1]
+         parameters$sigma <- log(pstart[2])
+         parameters$nu <- log(pstart[3])
+     }
+     mod <- new("exgaus", parameters = parameters, fixed = fixed, 
+         x = x, y = y, npar = npar)
+     mod
+ })
\end{Sinput}
\begin{Soutput}
[1] "exgaus"
\end{Soutput}
\end{Schunk}

>>>>

\paragraph{Starting values} Note also that start values for the
transition parameters are provided in this call using the
\code{trstart} argument.  At this time, the package does not provide
automatic starting values.  Starting values for parameters can be
provided using three arguments: \code{instart} for the parameters
relating to the prior probabilities, \code{trstart}, relating the
transition models, and \code{respstart} for the parameters of the
response models.  Note that the starting values for the initial and
transition models as well as of the multinomial logit response models
are interpreted as {\em probabilities}, and internally converted to
multinomial logit parameters.

\paragraph{Fitting the model and printing results} The \code{depmix}
function returns an object of class \code{depmix} which contains the
model specification, and not a fitted model.  Hence, the model needs
to be fitted by calling \code{fit}:
\begin{Schunk}
\begin{Sinput}
> fm <- fit(mod)
\end{Sinput}
\begin{Soutput}
iteration 0 logLik: -305.3323 
iteration 5 logLik: -305.3291 
iteration 10 logLik: -305.3234 
iteration 15 logLik: -305.3093 
iteration 20 logLik: -305.2689 
iteration 25 logLik: -305.1367 
iteration 30 logLik: -304.5522 
iteration 35 logLik: -291.1151 
iteration 40 logLik: -87.53257 
iteration 45 logLik: -84.34731 
iteration 50 logLik: -84.34178 
iteration 54 logLik: -84.34176 
\end{Soutput}
\end{Schunk}

The \code{fit}-function returns an object of class
\code{depmix.fitted} which extends the \code{depmix} class, adding
convergence information (and information about constraints if these
were applied, see below).  The class has \code{print} and
\code{summary} methods to see the results.  The \code{print} method
provides summary information on convergence, the log-likelihood and
the AIC and BIC values:
\begin{Schunk}
\begin{Sinput}
> fm
\end{Sinput}
\begin{Soutput}
Convergence info: Log likelihood converged to within tol. (relative change crit.) 
'log Lik.' -84.34176 (df=7)
AIC:  182.6835 
BIC:  211.275 
\end{Soutput}
\end{Schunk}

@@

The \code{summary} method of \code{fit}ted models provides the
parameter estimates, first for the prior probabilities model, second
for the transition models, and third for the response models.
\begin{Schunk}
\begin{Sinput}
> summary(fm)
\end{Sinput}
\begin{Soutput}
Initial state probabilties model 
Model of type multinomial, formula: ~1
Coefficients: 
     [,1]      [,2]
[1,]    0 -11.08968
Probalities at zero values of the covariates.
0.9999847 1.526887e-05

Transition model for state (component) 1 
Model of type multinomial, formula: ~1
Coefficients: 
[1] 0.9162498 0.0837502

Transition model for state (component) 2 
Model of type multinomial, formula: ~1
Coefficients: 
[1] 0.1053408 0.8946592

@@

Response model for response 1 
Model of type gaussian, formula: rt ~ 1
Coefficients: 
[1] 6.385491
sd  0.2439384

@@

Response model for response 1 
Model of type gaussian, formula: rt ~ 1
Coefficients: 
[1] 5.511149
sd  0.1926048 
\end{Soutput}
\end{Schunk}

@@

\begin{figure}[htbp]
\begin{center}
	\includegraphics{graphs/baldist.pdf}
	\caption{Balance scale item; this is a so-called distance item.}
	\label{fig:balance}  
\end{center}
\end{figure}

@@

Loading required package: Rsolnp (version 0.3)
    1 fx=   0.000000e+00 upsi=  6.7e+00 b2n= -1.0e+00 umi=  0.0e+00 nr   1 si-1
    2 fx=   2.640133e+02 upsi=  0.0e+00 b2n=  2.1e+01 umi=  0.0e+00 nr   1 si-1
    3 fx=   2.636388e+02 upsi=  0.0e+00 b2n=  5.4e+01 umi=  0.0e+00 nr   1 si-1
    4 fx=   2.587645e+02 upsi=  0.0e+00 b2n=  2.7e+02 umi=  0.0e+00 nr   1 si-1
    5 fx=   2.528471e+02 upsi=  0.0e+00 b2n=  1.2e+02 umi=  0.0e+00 nr   1 si-1
    6 fx=   2.523984e+02 upsi=  0.0e+00 b2n=  9.6e+01 umi=  0.0e+00 nr   1 si-1
    7 fx=   2.523907e+02 upsi=  0.0e+00 b2n=  9.5e+01 umi=  0.0e+00 nr   1 si-1
    8 fx=   2.522057e+02 upsi=  0.0e+00 b2n=  9.2e+01 umi=  0.0e+00 nr   1 si-1
    9 fx=   2.515501e+02 upsi=  0.0e+00 b2n=  5.8e+01 umi=  0.0e+00 nr   1 si-1
   10 fx=   2.510856e+02 upsi=  0.0e+00 b2n=  5.2e+01 umi=  0.0e+00 nr   1 si-1
   11 fx=   2.507418e+02 upsi=  0.0e+00 b2n=  7.2e+01 umi=  0.0e+00 nr   1 si-1
   12 fx=   2.503033e+02 upsi=  0.0e+00 b2n=  1.9e+01 umi=  0.0e+00 nr   1 si-1
   13 fx=   2.502725e+02 upsi=  0.0e+00 b2n=  9.8e-01 umi=  0.0e+00 nr   1 si-1
   14 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.5e-01 umi=  0.0e+00 nr   1 si-1
   15 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  3.1e-02 umi=  0.0e+00 nr   1 si-1
   16 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  4.1e-02 umi=  0.0e+00 nr   1 si-1
   17 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  6.3e-02 umi=  0.0e+00 nr   1 si-1
   18 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  2.9e-02 umi=  0.0e+00 nr   1 si-1
   19 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  4.1e-03 umi=  0.0e+00 nr   1 si-1
   20 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.4e-04 umi=  0.0e+00 nr   1 si-1
   21 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  4.3e-03 umi=  0.0e+00 nr   1 si-1
   22 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  9.4e-02 umi=  0.0e+00 nr   1 si-1
   23 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.1e-01 umi=  0.0e+00 nr   1 si-1
   24 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.1e-01 umi=  0.0e+00 nr   1 si-1
   25 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  2.7e-02 umi=  0.0e+00 nr   1 si-1
   26 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  4.6e-02 umi=  0.0e+00 nr   1 si-1
   27 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  2.0e-03 umi=  0.0e+00 nr   1 si-1
   28 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  2.0e-02 umi=  0.0e+00 nr   1 si-1
   29 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  4.9e-02 umi=  0.0e+00 nr   1 si-1
   30 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   31 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   32 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   33 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   34 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   35 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.2e-01 umi=  0.0e+00 nr   1 si-1
   36 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.1e-01 umi=  0.0e+00 nr   1 si-1
   37 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.1e-01 umi=  0.0e+00 nr   1 si-1
   38 fx=   2.502723e+02 upsi=  0.0e+00 b2n=  1.5e-02 umi=  0.0e+00 nr   1 si-1

@@

 termination reason:
 KT-conditions satisfied, no further correction computed
 evaluations of f                          149
 evaluations of grad f                      38
 evaluations of constraints                150
 evaluations of grads of constraints         0
 final scaling of objective           2.578752e+00
 norm of grad(f)                      5.469760e-01
 lagrangian violation                 3.850303e-06
 feasibility violation                0.000000e+00
 dual feasibility violation           0.000000e+00
 optimizer runtime sec's              1.020294e+02

@@

The constructor method return an object of class \code{exgaus}, and is
defined as follows:
\begin{Schunk}
\begin{Sinput}
> setGeneric("exgaus", function(y, pstart = NULL, fixed = NULL, 
+     ...) standardGeneric("exgaus"))
\end{Sinput}
\begin{Soutput}
[1] "exgaus"
\end{Soutput}
\begin{Sinput}
> require(gamlss)
\end{Sinput}
\begin{Soutput}
 **********   GAMLSS Version 3.0-1 ********** 
For more on GAMLSS look at http://www.gamlss.com/ 
Type gamlssNews() to see new features/changes/bug fixes.
\end{Soutput}
\begin{Sinput}
> require(gamlss.dist)
> setMethod("exgaus", signature(y = "ANY"), function(y, pstart = NULL, 
+     fixed = NULL, ...) {
+     y <- matrix(y, length(y))
+     x <- matrix(1)
+     parameters <- list()
+     npar <- 3
+     if (is.null(fixed)) 
+         fixed <- as.logical(rep(0, npar))
+     if (!is.null(pstart)) {
+         if (length(pstart) != npar) 
+             stop("length of 'pstart' must be ", npar)
+         parameters$mu <- pstart[1]
+         parameters$sigma <- log(pstart[2])
+         parameters$nu <- log(pstart[3])
+     }
+     mod <- new("exgaus", parameters = parameters, fixed = fixed, 
+         x = x, y = y, npar = npar)
+     mod
+ })
\end{Sinput}
\begin{Soutput}
[1] "exgaus"
\end{Soutput}
\end{Schunk}
